---
title: "assignment2bird"
author: "Stephen Shannon"
date: "February 23, 2021"
output:
  word_document: default
  html_document: default
---

##Loading Data in and libraries
```{r}
library(readxl)
library(vegan)
birddata <- read_excel("birddata.xlsx")
birddata
```
##Calculating True diversity by taking exponentiating Shannon's index
```{r}
#thankfully we just opened a library which has a function for us, R does not like to take the natural log of 0, which is a place holder in our syndicate vector to allow for a Completely randomized pair design.
syndicate <- (birddata$Syndicate)
springfield <- (birddata$Springfield)
syndicate_shan <- exp(diversity(syndicate, index="shannon"))
springfield_shan <- exp(diversity(springfield, index="shannon"))
diff <- abs(syndicate_shan - springfield_shan)
print(diff)
```
We calculate the absolute difference between exp(shannon's index) for the two groups to be 2.985. The exponent of our Shannon's index value will be our test statistic in the upcoming Completely randomized paired design.

##Completely randomized test design

We wish to know if the site location has an affect exp(Shannon's index). We will thus setup the following two tailed hypothesis test: 

H0: Tempature does not affect bulb burnout time. [exp(Shannon's index of syndicate) - exp(Shannon's index of springfield) = 0]
HA: Tempature does affect bulb burnout time. [exp(Shannon's index of syndicate) - exp(Shannon's index of springfield) is not equal to  0]


In order to carry out this test, we would need to make 2^20 permutations. This would be a very large vector that most computers would not be able to handle, so instead we will estimate the P-value for the following test by running Monte Carlo simulation on our two data vectors.


```{r}

shannon <- function(data){
  val <- exp(diversity(data,index="shannon"))
  return(val)
}

shannon(syndicate)
```

```{r}
bird <- c(syndicate, springfield)
N <- 10000
res <- numeric(N)
for (i in 1:N)
  {index <- as.numeric(sample(40,size=20,replace=F))
    res[i] <- shannon(bird[index])-shannon(bird[-index])}

observed <- shannon(syndicate)-shannon(springfield)
pval <- (sum(abs(res) >= abs(observed))+1)/(N+1)
round(pval,2)


```
We get a P-value of 0.12, which indicates the following: The probability of observing a value as extreme or more extreme than our observed difference in True Diversity between the syndicate and springfield sites is 0.12. Unfortunately, since P > 0.05, we do not have sufficient evidence to reject the null hypothesis. 

